---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "# "
)
```


# dplyr-cli

This a fork of a very nice tool written as a Rscript uses the `Rscript` executable to 
run dplyr commands on CSV files in the terminal. This version, however, changes 
the code so it reads any filetype supported by `data.table::fread()` and manipulates
the data as a `data.table` object. 

`dplyr-cli` makes use of the terminal pipe `|` instead of the magrittr pipe (`%>%`)
to run sequences of commands.

```
cat mtcars.csv | group_by cyl | summarise "mpg = mean(mpg)" | kable
#> | cyl|      mpg|
#> |---:|--------:|
#> |   4| 26.66364|
#> |   6| 19.74286|
#> |   8| 15.10000|
```

## Motivation

This modification of the original code was a result of necessity while working with genotype data. It 
is by no means the quickest way of doing this type of manipulation but it reduces learning curve while
keeping the process relatively quick due to `data.table` use.


## What dplyr commands are supported?

Any command of the form:

* `dplyr::verb(.data, code)`
* `dplyr::*_join(.data, .rhs)`

Currently two extra commands are supported which are not part of `dplyr`.

* `csv` performs no dplyr command, but only outputs the input data as CSV to stdout
* `kable` performs no dplyr command, but only outputs the input data as a
  `knitr::kable()` formatted string to stdout


## Limitations

* Only tested under 'bash'  and 'zsh' on OSX and Linux. YMMV.
* Every command runs in a separate R session.
* When using special shell characters such as `()`, you'll have to quote 
  your code arguments.  Some shells will require more quoting than others.
* "joins" (such as `left_join`) do not currently let you specify the `by` argument, 
  so there must be columns in common to both dataset

## Usage

```{sh}
dplyr --help
```

## History


#### v0.1.0 2020-04-20

* Initial release

#### v0.1.1 2020-04-21

* Switch to 'Rscript' for easier install for users
* rename 'dplyr.sh' to just 'dplyr'

#### v0.1.2 2020-04-21

* Support for joins e.g. `left_join`

#### v0.1.3 2020-04-22

* More robust tmpdir handling

#### v0.1.4 2022-01-23

* Fix handling for latest `read_csv()`.  Fixes #9

#### v0.2.0 2020-06-02

* Converted to use fread and data.table

## Contributors

* [coolbutuseless](https://github.com/coolbutuseless) - original author
* [aborusso](https://github.com/aborruso) - documentation


## Installation

Because this script straddles a great divide between R and the shell, you need 
to ensure both are set up correctly for this to work.

1. ~Install R packages~. This was also edited to perform installation of dependencies in first run. So in the very first run, expect it not to work. I still have to figure out a way of echoing meaninful warnings in this case.
2. Clone this repo and put `dplyr` in your path


#### Clone this repo and put `dplyr` in your path


You'll then need to download the shell script from this repository and put `dplyr`
somewhere in your path.

```
git clone https://github.com/coolbutuseless/dplyr-cli
cp dplyr-cli/dplyr ./somewhere/in/your/search/path
```


# Example data

Put an example CSV file on the filesystem. Note: This CSV file is now included as 
`mtcars.csv` as part of this git repository, as is a second CSV file for 
demonstrating joins - `cyl.csv`

```{r}
write.csv(mtcars, "mtcars.csv", row.names = FALSE)
```

# Example 1 - Basic Usage


```{sh}
# cat contents of input CSV into dplyr-cli.  
# Use '-c' to output CSV if this is the final step
cat mtcars.csv | dplyr filter -c "mpg == 21"
```


```{sh}
# Put quotes around any commands which contain special characters like <>()
cat mtcars.csv | dplyr filter -c "mpg < 11"
```


```{sh}
# Combine dplyr commands with shell 'head' command
dplyr select --file mtcars.csv -c cyl | head -n 6
```


# Example 2 - Simple piping of commands (with shell pipe, not magrittr pipe)

```{sh}
cat mtcars.csv | \
   dplyr mutate "cyl2 = 2 * cyl"  | \
   dplyr filter "cyl == 8" | \
   dplyr kable
```


# Example 3 - set up some aliases for convenience


```{sh}
alias mutate="dplyr mutate"
alias filter="dplyr filter"
alias select="dplyr select"
alias summarise="dplyr summarise"
alias group_by="dplyr group_by"
alias ungroup="dplyr ungroup"
alias count="dplyr count"
alias arrange="dplyr arrange"
alias kable="dplyr kable"


cat mtcars.csv | group_by cyl | summarise "mpg = mean(mpg)" | kable
```


# Example 4 - joins

Limitations:

* first argument after a join command must be an existing file 
* You can't yet specify a `by` argument for a join, so there must be a column in 
  common to join by
  
  
```{sh}
cat cyl.csv
```


```{sh}
cat mtcars.csv | dplyr inner_join cyl.csv | dplyr kable
```


## Security warning

`dplyr-cli` uses `eval(parse(text = ...))` on user input.  Do not expose this 
program to the internet or random users under any circumstances.


## Inspirations

* [xsv](https://github.com/BurntSushi/xsv) - a fast CSV command line toolkit 
  written in Rust
* [jq](https://stedolan.github.io/jq/) - a command line JSON processor.
* [miller](http://johnkerl.org/miller/doc/)
